---
title: "A Systematic Literature Review of Undergraduate Data Science Education Research"
format:
  jasa-pdf:
    keep-tex: true  
    journal:
      blinded: false
  jasa-html: default

date: last-modified
author:
  - name: Mine Dogucu
    acknowledgements: | 
      Dogucu has been supported by NSF IIS award #2123366. Dogucu completed an earlier part of this work in Department of Statistical Science at University College London.
    affiliations:
      - name: University of California, Irvine
        department: Department of Statistics
  - name: Sinem Demirci
    acknowledgements: | 
      Demirci has been supported by the Scientific and Technological Research Council of Türkiye. Demirci completed an earlier part of this work in Department of Statistical Science at University College London.
    affiliations: 
      - name: California Polytechnic State University
        department: Statistics Department
  - name: Harry Bendekgey
    acknowledgements: | 
      Bendekgey has been supported by the HPI Research Center in Machine Learning and Data Science at UC Irvine.
    affiliations:
      - name: University of California, Irvine
        department: Department of Computer Science
  - name: Federica Zoe Ricci
    acknowledgements: | 
      Ricci has been supported by the HPI Research Center in Machine Learning and Data Science at UC Irvine.
    affiliations:
      - name: University of California, Irvine
        department: Department of Statistics
  - name: Catalina M. Medina
    acknowledgements: | 
      Medina has been supported by NSF IIS award #2123366.
    affiliations:
      - name: University of California, Irvine
        department: Department of Statistics
abstract: |
  The presence of data science has been profound in the scientific community in almost every discipline. An important part of the data science education expansion has been at the undergraduate level. We conducted a systematic literature review to (1) portray current evidence and knowledge gaps in self-proclaimed undergraduate data science education research and (2) inform policymakers and the data science education community about what educators may encounter when searching for literature using the general keyword 'data science education.' While open-access publications that target a broader audience of data science educators and include multiple examples of data science programs and courses are a strength, significant knowledge gaps remain. The undergraduate data science literature that we identified often lacks empirical data, research questions and reproducibility. Certain disciplines are less visible. We recommend that we should (1) cherish data science as an interdisciplinary field; (2) adopt a consistent set of keywords/terminology to ensure data science education literature is easily identifiable; (3) prioritize investments in empirical studies.
  
keywords:
  - data science curriculum
  - data science programs
  - data science courses
  - educational technology
  - open access
bibliography: references.bib
---

```{=html}
<!--Blind links manually
\newcommand{\github}[1]{LINK HIDDEN FOR REVIEW}
\newcommand{\osf}[1]{LINK HIDDEN FOR REVIEW}
\newcommand{\ecots}[1]{CONFERENCE NAME HIDDEN FOR REVIEW}
-->
```
\newcommand{\github}[1]{https://github.com/mdogucu/comp-data-sci}
\newcommand{\osf}[1]{https://osf.io/b3u7y/}
\newcommand{\ecots}[1]{Electronic Conference on Teaching Statistics 2024}

```{=html}
<!--
\textcolor{red}{Mine}
\textcolor{purple}{Sinem}
\textcolor{brown}{Harry}
\textcolor{orange}{Federica}
\textcolor{green}{Catalina}
-->
```
## Introduction {#sec-introduction}

```{r}
#| child: 01-introduction.qmd
```

## Data Collection and Analysis {#sec-methodology}

```{r}
#| child: 02-methodology.qmd
```

## Results {#sec-results}

```{r}
#| child: 03-results.qmd
```

## Discussion {#sec-discussion}

```{r}
#| child: 04-discussion.qmd
```

## Data Availability Statement {.unnumbered}

The data on all the publications and the associated codebook for the variables are publicly available in a GitHub repository at \github{} and an OSF project at \osf{}.