---
title: "Data Cleaning - All Reviews"
format: html
editor: source
---

## Cleaning

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(here))
```

### Read files files from 04/04 merged_full

The documents we did not keep from the preliminary analysis need to be collected.

```{r}
prelim_reviews_og <- read_csv(
  here("data", "04-preliminary-analysis", "04_merged_reviews", "merged_full.csv"),
  show_col_types = FALSE, 
  col_types = cols(.default = col_character(), keep = col_logical())
)

prelim_docs_not_kept <- prelim_reviews_og %>% 
  filter(keep == FALSE) %>% 
  mutate(prelim_review = TRUE)
```

### Read files from 05/02

We also need all of the documents that made it past the preliminary analysis

```{r}
in_depth_folder_path <- here("data", "05-in-depth-analysis", "02-combined-reviews")
csv_names <- list.files(in_depth_folder_path)
combined_csv_paths <- paste0(in_depth_folder_path, "/", csv_names)

in_depth_reviews_og <- read_csv(
  combined_csv_paths[1], 
  show_col_types = FALSE, 
  col_types = cols(.default = col_character(), keep = col_logical())
)

for(path in combined_csv_paths[-1]) {
  in_depth_reviews_og <- bind_rows(
    in_depth_reviews_og,
    read_csv(
      path, 
      show_col_types = FALSE, 
      col_types = cols(.default = col_character(), keep = col_logical())
    )
  )
}

in_depth_reviews <- in_depth_reviews_og %>% 
  mutate(prelim_review = FALSE)
```

### Merge these files

We need to merge both the preliminary not kept and in-depth analysis documents.
These data frames will not have the same columns.
The column `document_type` should be dropped because we changed replaced this with `new_doc_type` in the in-depth reviews.

```{r}
all_reviews_og <- full_join(
  prelim_docs_not_kept, 
  in_depth_reviews,
  by = join_by(
    document_title, doi, document_link, authors, publication_year, abstract, database, open_access, reviewer_1, reviewer_2, keep, why_not_keep, prelim_review
  )
) 
```

### Clean them

For all docs with `keep = FALSE` set the following values to NA: `open_access`, `new_doc_type`, `journal`, `conference`, `book`, `affiliation_country`, `content_area`, `research_question`, `data_collection_type`, `main_field`, `big_pic_notes`, `interesting_facts`, `keywords`

```{r}
all_reviews_kept <- all_reviews_og %>% 
  filter(keep == TRUE)

all_reviews_not_kept <- all_reviews_og %>% 
  filter(keep == FALSE) %>% 
  mutate(
    open_access = NA,
    new_doc_type = NA,
    journal = NA,
    conference = NA,
    book = NA,
    affiliation_country = NA,
    content_area = NA,
    research_question = NA,
    data_collection_type = NA,
    main_field = NA,
    big_pic_notes = NA,
    interesting_facts = NA,
    keywords = NA,
  )

all_reviews <- bind_rows(all_reviews_kept, all_reviews_not_kept)
```

One document was categorized as a conference article but it is actually a journal article (in the first arXiv version, the below line incorrectly set the journal venue to "TRUE" but left the document type as "conference article")

```{r}
doc3 <- which(all_reviews$our_id == "Ceccucci_2014_effectiveness")
all_reviews$new_doc_type[doc3] <- "journal article"
```

Duplicates named differently for `why_not_keep`, so replace:

-   "duplicate" and "duplicated paper" with "duplicate paper"
-   "mooc" and "moocs" with "Massive Open Online Course"
-   "not about data science education" and "not data science education" with "not on data science education"
-   "not about data science education (about using data science to analyze education-related data)" with "not on data science education (about using data science to analyze education-related data)"

```{r clean-why_not_keep}
all_reviews <- all_reviews %>% 
  select(-c(document_type)) %>% 
  mutate(why_not_keep = str_to_lower(why_not_keep)) %>%
  mutate(why_not_keep = case_when(
    why_not_keep %in% c("duplicate", "duplicated paper") ~ "duplicate paper",
    why_not_keep %in% c("mooc", "moocs") ~ "Massive Open Online Course",
    why_not_keep %in% c("not about data science education", "not data science education") ~ "not on data science education",
    why_not_keep == "not about data science education (about using data science to analyze education-related data)" ~ "not on data science education (about using data science to analyze education-related data)",
    .default = why_not_keep
  )) 
```

Invalid country name is `affiliation_country`: replace "United States, United States" with "United States of America, United States of America".

```{r clean-affiliation_country}
all_reviews <- all_reviews %>% 
  mutate(affiliation_country = case_when(
    affiliation_country == "United States, United States" ~ "United States of America, United States of America",
    affiliation_country == "Netherlands, Spain, England, Norway" ~ "Netherlands, Spain, United Kingdom, Norway",
    .default = affiliation_country
  ))
```

For docs with two content areas make sure they are listed alphabetically.

Invalid entry in content_area, replace: "current state of data science education" with "review of current state of data science education"

```{r clean-content_area}
listed_content_areas <- str_split(all_reviews$content_area, pattern = " and ")
sorted_listed_content_areas <- lapply(listed_content_areas, sort, na.last = TRUE)
rearranged_content_areas <- unlist(
  lapply(sorted_listed_content_areas, str_c, collapse = " and ")
)

all_reviews <- all_reviews %>% 
  mutate(content_area = rearranged_content_areas) %>% 
  mutate(content_area = if_else(
    content_area == "current state of data science education", 
    true = "review of current state of data science education",
    false = content_area
  ))
```

Duplicates names differently in `data_collection_type`, replace: "quantitative" with "quant" (also fix the single row with NA `data_collection_type` to "no data").

Also, "Cuadrado_2021_classification" was missclassified as `data_collection_type` being "no data" but it should be "mixed".

```{r clean-data_collection_type}
all_reviews <- all_reviews %>% 
  mutate(data_collection_type = case_when(
    data_collection_type == "quantitative" ~ "quant",
    is.na(data_collection_type) & keep ~ "no data",
    .default = data_collection_type
  ))

reclassified_data_paper <- which(all_reviews$our_id == "Cuadrado_2021_classification")
all_reviews$data_collection_type[reclassified_data_paper] <- "mixed"
```

Clean `keywords` by making everything lowercase and alphabetically ordered, replace:

-   "data science\neducation" with "data science education"
-   "program accreditation\n" with "program accreditation"
-   "\ndata science" with "data science"
-   "instructional\ndesign" with "instructional design"
-   ";" with "," (because of "Shapiro_2020_reshape") entry
-   " ," with ","
-   `NA` to "not_specified" for "Davis_2020_ethics" specifically

```{r}
listed_keywords <- str_split(all_reviews$keywords, pattern = ", ")
sorted_listed_keywords <- lapply(listed_keywords, sort, na.last = TRUE)
rearranged_keywords <- unlist(
  lapply(sorted_listed_keywords, str_c, collapse = ", ")
)

all_reviews <- all_reviews %>% 
  mutate(keywords = str_to_lower(rearranged_keywords)) %>% 
  mutate(keywords = str_replace(
    keywords,
    pattern = "data science\neducation", 
    replacement = "data science education"
  )) %>% 
  mutate(keywords = str_replace(
    keywords,
    pattern = "program accreditation\n", 
    replacement = "program accreditation"
  )) %>% 
  mutate(keywords = str_replace(
    keywords,
    pattern = "\ndata science", 
    replacement = "data science"
  )) %>% 
  mutate(keywords = str_replace(
    keywords,
    pattern = "instructional\ndesign", 
    replacement = "instructional design"
  )) %>% 
  mutate(keywords = str_replace(
    keywords,
    pattern = ";", 
    replacement = ","
  )) %>%
  mutate(keywords = str_replace(
    keywords,
    pattern = " ,", 
    replacement = ","
  ))
  
no_keywords_index <- which(all_reviews$our_id == "Davis_2020_ethics")

all_reviews[no_keywords_index, "keywords"] <- "not specified"
```

The format of papers that were excluded for a reason other than their format (as reflected by their `why_not_keep` variable) was re-checked. If the format was found to be among the ones we decided to exclude (e.g. editorial material), the reason for their exclusion is changed to their format:

```{r}

posters_doi <- c("10.1145/2992154.2996788", 
                 "10.1145/3488042.3490509",
                 "10.1145/3488042.3489966")
posters <- (!is.na(all_reviews$doi) & all_reviews$doi %in% posters_doi) 
all_reviews$why_not_keep[posters] <- "poster"

not_english_doi <- c("10.1541/ieejeiss.142.291")
not_english <- (!is.na(all_reviews$doi) & all_reviews$doi %in% not_english_doi) 
all_reviews$why_not_keep[not_english] <- "not in english"

edit_material_doi <- c("10.1111/bjet.13258")
edit_materials <- (!is.na(all_reviews$doi) & all_reviews$doi %in% edit_material_doi) 
all_reviews$why_not_keep[edit_materials] <- "editorial material"

# panel_doi <- c("10.1002/pra2.84")
# panels <- (!is.na(all_reviews$doi) & all_reviews$doi %in% panel_doi) 
# all_reviews$why_not_keep[panels] <- "panel"



```

## Update Sep 2024 to publication years

```{r}
updated_year_csv <- c()
for(reviewer in c('catalina', 'federica', 'harry', 'mine', 'sinem')){
  path <- here("data", "06-data-cleaning", "add-year-first-published", paste(reviewer, '.csv', sep=''))
  updated_year_csv <- rbind(updated_year_csv, read_csv(path, 
                                                      show_col_types = FALSE, 
                                                      col_types = cols(.default = col_character())))
}
updated_year_csv <- rename(updated_year_csv, 
                           document_link_ref = document_link, 
                           reviewer_1_ref = reviewer_1, 
                           reviewer_2_ref = reviewer_2)


disagreements <- (updated_year_csv |> group_by(our_id) |> mutate(n_decisions = n_distinct(first_published)) |> filter(n_decisions > 1))$our_id

stopifnot(length(disagreements) == 0)

updated_year_csv <- updated_year_csv |> group_by(our_id) |> summarise(first_published = min(first_published))

all_reviews <- merge(all_reviews, updated_year_csv, by='our_id', all.x = TRUE)
```

## Change RQ and data classification of Banadaki_2022_enabling_data based on Nov 2024 Meeting (Harry + Federica)

```{r}
doc <- which(all_reviews$our_id == "Banadaki_2022_enabling_data")
all_reviews$research_question[doc] <- "TRUE"
all_reviews$data_collection_type[doc] <- "qual"
```

## Replace main_field with multiple variables: audience_discipline, observational_unit, department and source_discipline

Remove `main_field` variable from `all_reviews`:

```{r}
all_reviews <- all_reviews |> 
  dplyr::select(-main_field)
```

Load file with `audience_discipline`, `observational_unit` and `department`

```{r}
recoded_main_field_all <- read_csv(
  here("data", 
       "06-data-cleaning", 
       "recode-main-field",
       "recoded_main_field_all.csv"), 
  show_col_types = FALSE, 
  col_types = cols(.default = col_character())
)
```

Load file with `source_discipline` for conferences and for journals + books.

```{r}
source_discipline_conferences <- read_csv(
  here("data", 
       "06-data-cleaning", 
       "code-source-discipline",
       "source_discipline_conferences.csv"), 
  show_col_types = FALSE, 
  col_types = cols(.default = col_character())) |> 
  select(-doc_type)

source_discipline_journals_books <- read_csv(
  here("data", 
       "06-data-cleaning", 
       "code-source-discipline",
       "source_discipline_journals_books.csv"), 
  show_col_types = FALSE, 
  col_types = cols(.default = col_character())) |> 
  select(-doc_type)


```

Add variable `source` to `all_reviews`

```{r}
all_reviews <- all_reviews |> 
  mutate(
    source = case_when(
      !is.na(journal) ~ journal,
      !is.na(conference) ~ conference,
      !is.na(book) ~ book
    )
  )
```


Join new variables

```{r}

all_source_discipline <- bind_rows(
  source_discipline_conferences,
  source_discipline_journals_books
)
  
all_reviews <- all_reviews |> 
  left_join(
    recoded_main_field_all,
    by = "our_id"
  ) |> 
  left_join(
    all_source_discipline,
    by = c("source")
  )  
```

## Write the final clean merged csv to in the data folder

```{r}
# Reorder the columns. Also drops the old publication_year column that has been replaced by first_published

col_order <- c("our_id", "document_title", "doi",
               "document_link", "authors", "abstract",
               "database", "reviewer_1", "reviewer_2", 
               "keep", "why_not_keep",  "prelim_review",
               "open_access", "first_published",
               "new_doc_type", 
               "journal", "conference", "book",
               "source",
               "affiliation_country", 
               "content_area", "research_question",
               "data_collection_type", 
               "audience_discipline",
               "reason_audience_discipline",
               "observational_unit", 
               "department", "reason_department", 
               "source_discipline",
               "reason_source_discipline",
               "big_pic_notes", 
               "interesting_facts", "keywords")
all_reviews <- all_reviews[, col_order]
```

```{r}
write_csv(all_reviews, here("data", "all_reviews.csv"))
```
